\sysfull

### 2017-08-03 paper/compiler.tex
committing partial edit before quick phone call

A unique feature of \sys is a compiler support for \emph{automatic task generation}. \sys literally eliminates the programmer's burden of writing a task-based program from scratch. In essence, the \sys compiler obtains a plain C code as an input and generates a binary file that is rewritten in task-based programming style that \sys executes.

The compiler consists of two {\em passes} that runs on LLVM Intermediate Representation (IR)\todo{Which version?}{Kiwan}---an intermediate language generated by Clang~\cite{clang_website}:
%
\begin{enumerate}
	\item \textbf{Analysis Pass}, which analyzes the code structure and decides where to put task boundaries;
	\item \textbf{Transform Pass}, which converts the code into task-style program and fills it with necessary \sys keywords so that it can run on \sys.
\end{enumerate}
%
We shall now proceed with describing these two parts in detail.

\subsection{Compiler Analysis Pass}
\label{sec:compiler_analysis_pass}

In the first pass \sys compiler decides where to put task boundaries. By definition, a \sys task should adhere to the following correctness constraints:

\begin{enumerate}[label={\bf C\arabic*:}]
\item{No task should ever use up more energy than what the capacitor of the system can provide;} 
\item{A task should be single-entry.}\todo{Define single entry}{Kiwan}
\end{enumerate}
%
Also for the system to be efficient, the compiler should optimize the following points:
%
\begin{enumerate}[label={\bf O\arabic*:}]
\item{Tasks should be roughly the same size for efficient coalescing;}
\item{The number of protected variables should be minimized;}
\item{Task boundary should be visited as least as possible due to overhead on boundaries.}\todo{Define overhead on boundaries}{Kiwan}
\end{enumerate}

Analysis pass tries to meet constraints \{C1, C2\} and \{O1, O2, O3\} by (i) \emph{graph building}, (ii) \emph{loop analysis}, (iii) \emph{task preparation}, and (iv) \emph{single entry assurance} procedures described below.

@@

\textbf{Graph Building:} starting from a plain C code (see example in Figure~\ref{fig:comp_1}) \sys compiler first builds a control flow graph where (i) each vertices is a \emph{basic block} and (ii) each unidirectional edge is a \emph{branch from one basic block to another} (see example in Figure~\ref{fig:comp_2}). All function calls are inlined in the main function prior to building a graph.

Each vertex has a {\em price}, which represents the energy usage of the basic block. We used the number of LLVM IRs inside a vertex as the energy use metric. When forming a task, each task should contain a basic block so that the sum of price of all blocks within each tasks should be roughly the same (viz. O1) and does not exceed given energy budget (viz. C1). For our implementation, the programmer should provide the energy budget in compile time defined in number of LLVM IR instructions that the hardware can execute safely in one energy cycle. This number must be found experimentally.

@@

\textbf{Loop Analysis:} After constructing the graph and assigning the information on a price and live variables to each vertex and edge, respectively, the compiler begins with analyzing the program loops. The compiler first identifies all loops and then multiplies the loop count by the price of a vertices within the loop. The reason is, if the loop is inside a task it will use up energy that is roughly multiplied by the loop count compared to code that is outside of a loop. If the resulting price of a loop is too large\todo{how do you decide on the acceptable loop price?}{Kiwan}, the compiler places boundary at the back edge of the loop, and reverts the change to the price of the vertices inside. It means that if the loop is too large to fit in one task, each iteration should be a separate task execution.

>>>>

\sys provides compiler support for \emph{automatic task generation} that simplifies programming writing task-based code by automatically translating C code into a task-based executable.  Translating C code into task-based code automatically requires identifying program points to use as task boundaries, defining a set of task functions that begin and end at those boundaries, ensuring that all existing control-flow paths are either contained within a task, or are made explicit in inter-task control-flow, and identifying and instrumenting accesses to task-shared variables that need to be protected by \sys at runtime.  The \sys compiler first analyzes the control-flow graph (CFG) and, using a rule-based heuristic, determines at which points to place a task boundary.  After identifying where to put task boundaries, the \sys compiler transforms the imperative code into tasks, instrumenting protected memory accesses.

%The compiler consists of two {\em passes} that runs on LLVM Intermediate Representation (IR)\todo{Which version?}{Kiwan}---an intermediate language generated by Clang~\cite{clang_website}:
%
%\begin{enumerate}
%	\item \textbf{Analysis Pass}, which analyzes the code structure and decides where to put task boundaries;
%	\item \textbf{Transform Pass}, which converts the code into task-style program and fills it with necessary \sys keywords so that it can run on \sys.
%\end{enumerate}
%
%We shall now proceed with describing these two parts in detail.

\subsection{Task Boundary Identification Analysis}
\label{sec:compiler_analysis_pass}

\todo{Define single entry}{Kiwan}
\TODO{Brandon: the term "single-entry" is well-known in the community to mean that there is a single control-flow edge entering the task. Thinking about this a bit, is there a simple reason why we cannot have multiple-entry tasks? A totally plausible task graph could be like A->B B->C C->B C->A; here, B has two predecessors.  Is this fundamental?  Or an analysis convenience? I've left it as "single-entry" for now}

\sys analyzes a program to determine where to put task boundaries to divide the
program into tasks. A task is a single-entry, multiple-exit region of the CFG
and in a correct task decomposition, no task should ever consume more energy
than the device's energy buffer can store.  There are several factors that
determine whether tasks will execute efficiently. A task decomposition should
minimize the number of variables that are accessed by multiple tasks and, 
consequently, must be protected by \sys's paging mechanism.  Tasks should
be roughly the same size, simplifying \sys's dynamic reasoning about coalescing
by eliminating the need to consider variable task cost at run time.  To minimize
overhead introduced at task boundaries (i.e., commit), a task decomposition should 
maximize task size, which minimizes the frequency of task transitions.

%\begin{enumerate}[label={\bf C\arabic*:}]
%\item{No task should ever use up more energy than what the capacitor of the system can provide;} \TODO{We do not (cannot, actually) guarantee this, so I think we can, at best, say that this is an optimization constraint.  What do you think?}
%\item{A task should be single-entry.}\todo{Define single entry}{Kiwan} 
%\end{enumerate}
%
%For a task-decomposition to be efficient, the compiler should optimize the following points:
%
%\begin{enumerate}[label={\bf O\arabic*:}]
%\item{Tasks should be roughly the same size for efficient coalescing;}
%\item{The number of protected variables should be minimized;}
%\item{Task boundary should be visited as least as possible due to overhead on boundaries.}\todo{Define overhead on boundaries}{Kiwan}
%\end{enumerate}

\sys meets these correctness and efficiency requirements with a four-part
analysis composed of (i) \emph{graph construction}, (ii) \emph{loop analysis},
(iii) \emph{task preparation}, and (iv) \emph{single entry assurance}.

@@

\textbf{Graph Construction:} starting from a plain C code (see example in Figure~\ref{fig:comp_1}) \sys compiler first builds a control flow graph where (i) each vertices is a \emph{basic block} and (ii) each unidirectional edge is a \emph{branch from one basic block to another} (see example in Figure~\ref{fig:comp_2}). All function calls are inlined in the main function prior to building a graph.

Each vertex has a {\em cost}, which represents the energy usage of the basic block. We used the number of LLVM IRs inside a vertex as the energy use metric. When forming a task, each task should contain a basic block so that the sum of cost of all blocks within each tasks should be roughly the same (viz. O1) and does not exceed given energy budget (viz. C1). For our implementation, the programmer should provide the energy budget in compile time defined in number of LLVM IR instructions that the hardware can execute safely in one energy cycle. This number must be found experimentally.

@@

\textbf{Loop Analysis:} After constructing the graph and assigning the information on a cost and live variables to each vertex and edge, respectively, the compiler begins with analyzing the program loops. The compiler first identifies all loops and then multiplies the loop count by the cost of a vertices within the loop. The reason is, if the loop is inside a task it will use up energy that is roughly multiplied by the loop count compared to code that is outside of a loop. If the resulting cost of a loop is too large\todo{how do you decide on the acceptable loop cost?}{Kiwan}, the compiler places boundary at the back edge of the loop, and reverts the change to the cost of the vertices inside. It means that if the loop is too large to fit in one task, each iteration should be a separate task execution.
